Analyzing Information Transfer from Different Data Sources via Precise Bias-variance Asymptotics

### 2021-07-02 revision_journal/results.tex
Updates from Overleaf

%In Section \ref{sec_HPS_loss}, we give a rigorous analysis of the bias term  and variance term  in two different settings:
In case (i), we provide the exact asymptotics for both the bias and variance terms, together with almost sharp convergence rates. We will use these results to illustrate the effects of the \emph{bias-variance tradeoff} and \emph{varying sample sizes}. In case (ii), we will prove an exact asymptotic variance limit, and employ the result to show the intricate effect of the covariate shift. On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$. %Here case (ii) is actually more general than (i)

\paragraph{Illustrative examples.}
First, we show that if $\|\beta_0\|^2 \gg d^2$, then the global minimizer $\hat a$ of the function $g(a)$ in \eqref{eq_mtl_A12} is close to $1$.
	The proof of Proposition \ref{lem_hat_v} will be given in Appendix \ref{app_iso_cov_prop}.

	\begin{proposition}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
Suppose Assumption \ref{assm_big1} and the above setting of random effect model hold. %in Example \ref{ex_same_cov} hold.
Suppose that for a constant $c_0>0$,
	\be\label{para_rel}
	 \|\beta_0\|^2 \ge p^{c_0}d^2 +p^{-1/2+c_0}\sigma^2.
	\ee
 Then we have that for any small constant $c>0$ and large constant $ C>0$,
	%Let $c$ be a sufficiently small fixed constant.
	%In the setting of Corollary \ref{cor_MTL_loss},
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}
	 \hat a =1+ \OO\left(\frac{d^2}{\|\beta_0\|^2} + p^{-1/4+c} \frac{d+\sigma}{\|\beta_0\|} + p^{-C}\right) \quad \text{w.h.p.}
	\ee
\end{proposition}
%Suppose every $\beta^{(i)}$ consists of two random components, one that is shared among all tasks and one that is task-specific.
%%Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
%More precisely, %we have
%$$\beta^{(i)}=\beta_0 +\wt \beta^{(i)},\quad i=1,2,\cdots, t,$$
%where $\beta_0$ denotes the shared component, %whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1}\kappa^2$,
%and $\wt \beta^{(i)}$ denotes the $i$-th
%%Let $\beta^{(i)}$ be equal to $\beta_0$ plus a
%task-specific component whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2$.

We emphasize that unlike Theorem \ref{cor_MTL_loss}, for Proposition \ref{lem_hat_v} we do not assume $\Sigma^{(1)}=\Sigma^{(2)}$ and the Gaussian distributions of the entries of $Z^{(1)}$ and $Z^{(2)}$. Combining Lemma \ref{lem_HPS_loss}, Theorem \ref{cor_MTL_loss} and Proposition \ref{lem_hat_v}, and using  that $\|\beta^{(1)}-\beta^{(2)}\|^2 = (2+\oo(1))d^2$ with high probability, the predication loss $L(\hat{\beta}_2^{\MTL}(\hat a)) $ is approximately equal to
$$\ell(n_1,n_2)= \sigma^2 \cal L_1(1) +   2d^2 \cal L_2(1)= \frac{p\sigma^2}{n_1+n_2-p} +  2d^2  \frac{n_1^2 (n_1+n_2-p)+p n_1n_2}{(n_1+n_2)^2(n_1+n_2-p)}, $$
with high probability, where in the second step we obtain $\cal L_1(1)$ and $\cal L_2(1)$ through direct calculations.
Now we compare $\ell(n_1,n_2)$ with \eqref{L_STL_simple01}. First, we notice that the variance term $\sigma^2 \cal L_1(1)$ is smaller than $L(\hat{\beta}_2^{\STL} )$, while the bias term always increases (because the bias in single-task linear regression is zero). This gives a \emph{bias-variance tradeoff} for the predication loss of the HPS estimator. Second, calculating the derivative with respect to $n_1$, it is not hard to see that the variance term always decreases as $n_1$ increases, while the bias term always increases with $n_1$. Hence how the performance of the HPS estimator changes with respect to $n_1$ also depends on an intricate bias-variance tradeoff.

\paragraph{Comparison to other estimators.}
    \begin{align}
    L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
    &+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2   \lambda^2 \bigtr{\Sigma^{(2)} [(X^{(1)})^\top X^{(1)}]^{-1}  }\nonumber\\
    & + \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  (1-\lambda)^2 \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_AVE_simple}
    \end{align}
    and
    \begin{align}
    L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_STL_simple}
    \end{align}
    %Moreover, there exists a constant $C_0>0$ such that with high probability,
    %\be\label{loss_large}
    %g(a) > g(0) \quad \text{for all $a$ such that }|a|\ge C_0.
    %\ee
    %with high probability for any small constant $\e>0$.

@@

%The previous section assumes that all tasks have the same sample size and feature vectors.
In this section, we consider the most general setting, where the feature covariates $X^{(1)}$ and $X^{(2)}$ of the two tasks have both different sample sizes and different population covariance matrices.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
In particular, the fact that population covariance matrices differ across tasks is often called ``covariate shift''. %which is characterized by the matrix $(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$.
In this section, we describe the exact asymptotic variance limit in the high-dimensional setting, while the bias limit is much more complicated, and we can only give an estimate on it.

\iffalse
Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p} \text{ and }
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
We consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \cite{WZR20}).
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where }
	\hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
\fi

%Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Compared to the results in Section \ref{sec_sizeratio}, the spectrum of $\hat{\Sigma}(a)^{-1}$ now not only depends on the sample sizes of both tasks, but also depends on the ``misalignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
%However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
To capture this misalignment quantitatively, we introduce the covariate shift matrix %(rescaled by $a$)
$$ M(a) \define a(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}.$$
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1(a)\ge \lambda_2(a)\ge \dots\ge \lambda_p(a)$ be the singular values of $M$ in descending order. By \eqref{assm3}, these singular values satisfy that
\begin{equation}\label{assm32}
|a| \tau \le  \lambda_p(a) \le \cdots \le \lambda_2(a)\le \lambda_1(a) \le |a|\tau^{-1}. %, \quad \max\left\{\pi_A^{(n)}([0,\tau]), \pi_B^{(n)}([0,\tau])\right\} \le 1 - \tau .
\end{equation}
%for some small constant $0<\tau<1$.
The main result of this section is the following theorem on the variance limit, which characterizes  the exact dependence of $L_{\var}(a)$ on the singular values of $M$.

\begin{theorem}\label{thm_main_RMT}
%	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
%	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
%	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
%	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
%	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:
Under Assumption \ref{assm_big1}, for any small constant $c>0$, there exists a high probability event $\Xi$, on which the following estimate holds for $L_{\var}(a)$ in \eqref{Lvar}:
%equation \eqref{eq_variance_2task} $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ (leaving out $\sigma^2$) satisfies the following estimate:
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{L_{\var}(a)- \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 M(a)^\top M(a) + a_2  }  }}
				\le \frac{(n_1+n_2)^{2/\varphi + c}}{p^{1/2} (n_1+n_2)^{1/2}}\cdot\frac{p \sigma^2}{n_1+ n_2}  ,
			\end{align}
			 uniformly in all $a\in \R$. Here $(a_1,a_2)$ is the solution of the following system of equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2}, \quad
				a_1 + \frac1{n_1 + n_2}  \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}

 With \eqref{assm32}, it is easy to see that
$$ \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 M(a)^\top M(a) + a_2  }  } \sim \frac{p \sigma^2}{n_1+ n_2}.$$
Hence the right-hand side of \eqref{lem_cov_shift_eq} is much smaller than this main term by a factor of $p^{-1/2} (n_1+n_2)^{-1/2+2/\varphi + c}$, which we believe to be sharp up to the $(n_1+n_2)^c$ factor. Lemma \ref{fact_tr} can be also regarded as a special case of Theorem \ref{thm_main_RMT}. %also extends  to the inverse of the sum of two sample covariance matrices.
To see this, when $M(a)=0$, we solve equation \eqref{eq_a12extra} to obtain that
$$a_1 = \frac{n_1}{n_1+n_2},\quad a_2 = \frac{n_2-p}{n_1+ n_2},$$
and plug them into \eqref{lem_cov_shift_eq}.  The proof of Theorem \ref{thm_main_RMT} is based on some recent developments in random matrix theory \cite{Anisotropic}, and will be given in Appendix \ref{appendix RMT} of the supplement \cite{MTL_suppl}.
%For general $A_1,A_2$ that are not equal to one, we can still apply our result by rescaling $X^{(1)}$ and $M$ with $A_1 / A_2$.
%We defer a proof sketch of Theorem \ref{thm_main_RMT} until the end of the section.
%This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.

\paragraph{Illustrative examples.}
To describe quantitatively the bias-variance tradeoff phenomenon and the effect of varying sample sizes on the predication loss, we consider a \emph{random-effect model}, which has been studied for  single-task linear regression and ridge regression (see e.g. \cite{dobriban2020wonder,dobriban2018high}).

 %{How does hard parameter sharing scale with sample sizes and covariate shift $M$?} One can see that the variance limit depends intricately on both tasks' samples sizes and covariate shift. Next, we illustrate how varying them impact the prediction loss.
%\begin{example}%[Sample sizes]
%\label{ex_sample_ratio}
	%We first consider the impact of varying sample sizes.
	%Consider the random-effect model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.
	%We again consider the random effect model in Example \ref{ex_same_cov} with $t=2$.
	Suppose each $\beta^{(i)}$ consists of two random components, one that is shared among the two tasks and one that is task-specific. More precisely, we let
$$\beta^{(i)}=\beta_0 +\wt \beta^{(i)},\quad i=1,2, $$
where $\beta_0$ denotes the shared component, %whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1}\kappa^2$,
and $\wt \beta^{(i)}$ denotes the $i$-th
%Let $\beta^{(i)}$ be equal to $\beta_0$ plus a
task-specific component whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2$.

@@

\paragraph{Comparison to other estimators.}

@@

\subsection{Model and covariate shift}

>>>>

In case (i), we provide the exact asymptotics for both the bias and variance terms, together with almost sharp convergence rates. We will use these results to illustrate the effects of the \emph{bias-variance tradeoff} and \emph{varying sample sizes}. In case (ii), we will prove an exact asymptotic variance limit, and employ the result to show the intricate effect of the covariate shift. On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$. %Here case (ii) is actually more general than (i)

@@

%The previous section assumes that all tasks have the same sample size and feature vectors.
We begin by considering the covariate shift setting where both tasks have the same linear model but different population covariance matrices.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
%In particular, the fact that population covariance matrices differ across tasks is often called ``covariate shift''. %which is characterized by the matrix $(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$.
We show the exact asymptotic limit of the excess risk of the HPS estimator in the high-dimensional setting.

%By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
%\begin{align}
%	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where }
%	\hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
%\end{align}
First, we consider the variance. Recall from equation \eqref{Lvar} that the variance equation is equal to $\sigma^2 \cdot \bigtr{\Sigma^{(2)} (\hat\Sigma(a))^{-1}}$.
In particular, the matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices.
Thus, the expectation of $\hat{\Sigma}(a)$ is equal to a mixture of both tasks' population covariance matrices, with mixing proportions determined by their sample sizes.
%To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
%\begin{align}
%	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
%	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
%	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
%\end{align}
%Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
%Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Intuitively, the spectrum of $\hat{\Sigma}(a)^{-1}$ now not only depends on the sample sizes of both tasks, but also depends on how well ``aligned'' $\Sigma^{(1)}$ and $\Sigma^{(2)}$ are.
%However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
To capture this alignment quantitatively, we introduce the covariate shift matrix %(rescaled by $a$)
$$ M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}.$$
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1 \ge \lambda_2 \ge \dots\ge \lambda_p $ be the singular values of $M$ in descending order. %By \eqref{assm3}, these singular values satisfy that
%\begin{equation}\label{assm32}
%|a| \tau \le  \lambda_p(a) \le \cdots \le \lambda_2(a)\le \lambda_1(a) \le |a|\tau^{-1}. %, \quad \max\left\{\pi_A^{(n)}([0,\tau]), \pi_B^{(n)}([0,\tau])\right\} \le 1 - \tau .
%\end{equation}
%for some small constant $0<\tau<1$.
Our first main result is the following theorem on the variance limit, which characterizes the exact dependence of $L_{\var}(a)$ on the singular values of $M$.

\begin{theorem}[Variance asymptotics under covariate shift]\label{thm_main_RMT}
%	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
%	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
%	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
%	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
%	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:
Under Assumption \ref{assm_big1} (recalling $\varphi > 4$), for any small constant $c>0$, there exists a high probability event $\Xi$, on which the following estimate holds for $L_{\var}(a)$ in \eqref{Lvar}:
%equation \eqref{eq_variance_2task} $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ (leaving out $\sigma^2$) satisfies the following estimate:
	\begin{align}\label{lem_cov_shift_eq}
		\bigabs{L_{\var}(a)- \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 a^2\cdot M^\top M + a_2  }  }}
		\le \frac{(n_1+n_2)^{\frac 2{\varphi} - \frac 1 2 + c}}{p^{1/2}}\cdot\frac{p \sigma^2}{n_1+ n_2},
	\end{align}
	uniformly in all $a\in \R$. Here $(\alpha_1, \alpha_2)$ is the solution of the following system of equations
	\begin{align}
		\alpha_1 + \alpha_2 = 1- \frac{p}{n_1 + n_2}, \quad
		\alpha_1 + \frac1{n_1 + n_2}  \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 \alpha_1 a^2}{\lambda_i^2 \alpha_1 a^2 + \alpha_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
	\end{align}
\end{theorem}
\todo{changed $a_1, a_2$ to $\alpha_1, \alpha_2$ since they conflict with $a$}

Equation \eqref{lem_cov_shift_eq} thus characterizes the variance of the HPS estimator, up to an error term on the order of $O\big(\frac{\sigma^2}{p^{-1/2}}\big)$ (recall that $\varphi > 4$, $c$ is an arbitrarily small constant, $\frac{p}{n_1+n_2} \in (2 {\tau}, \frac 2 {\tau})$).
With Assumption \ref{assm_big1},
$$ \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 M(a)^\top M(a) + a_2  }  } \sim \frac{p \sigma^2}{n_1+ n_2}.$$
Hence the right-hand side of \eqref{lem_cov_shift_eq} is much smaller than this main term by a factor of $p^{-1/2} (n_1+n_2)^{-1/2+2/\varphi + c}$, which we believe to be sharp up to the $(n_1+n_2)^c$ factor. Lemma \ref{fact_tr} can be also regarded as a special case of Theorem \ref{thm_main_RMT}. %also extends  to the inverse of the sum of two sample covariance matrices.
To see this, when $M(a)=0$, we solve equation \eqref{eq_a12extra} to obtain that
$$a_1 = \frac{n_1}{n_1+n_2},\quad a_2 = \frac{n_2-p}{n_1+ n_2},$$
and plug them into \eqref{lem_cov_shift_eq}.  The proof of Theorem \ref{thm_main_RMT} is based on some recent developments in random matrix theory \cite{Anisotropic}, and will be given in Appendix \ref{appendix RMT} of the supplement \cite{MTL_suppl}.
%For general $A_1,A_2$ that are not equal to one, we can still apply our result by rescaling $X^{(1)}$ and $M$ with $A_1 / A_2$.
%We defer a proof sketch of Theorem \ref{thm_main_RMT} until the end of the section.
%This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.

\paragraph{Illustrative examples.}
To describe quantitatively the bias-variance tradeoff phenomenon and the effect of varying sample sizes on the predication loss, we consider a \emph{random-effect model},

@@

\paragraph{Comparing HPS to other estimators.}

@@

\paragraph{Illustrative examples.}
Combining Lemma \ref{lem_HPS_loss}, Theorem \ref{cor_MTL_loss} and Proposition \ref{lem_hat_v}, and using  that $\|\beta^{(1)}-\beta^{(2)}\|^2 = (2+\oo(1))d^2$ with high probability, the predication loss $L(\hat{\beta}_2^{\MTL}(\hat a)) $ is approximately equal to
$$\ell(n_1,n_2)= \sigma^2 \cal L_1(1) +   2d^2 \cal L_2(1)= \frac{p\sigma^2}{n_1+n_2-p} +  2d^2  \frac{n_1^2 (n_1+n_2-p)+p n_1n_2}{(n_1+n_2)^2(n_1+n_2-p)}, $$
with high probability, where in the second step we obtain $\cal L_1(1)$ and $\cal L_2(1)$ through direct calculations.
Now we compare $\ell(n_1,n_2)$ with \eqref{L_STL_simple01}. First, we notice that the variance term $\sigma^2 \cal L_1(1)$ is smaller than $L(\hat{\beta}_2^{\STL} )$, while the bias term always increases (because the bias in single-task linear regression is zero). This gives a \emph{bias-variance tradeoff} for the predication loss of the HPS estimator. Second, calculating the derivative with respect to $n_1$, it is not hard to see that the variance term always decreases as $n_1$ increases, while the bias term always increases with $n_1$. Hence how the performance of the HPS estimator changes with respect to $n_1$ also depends on an intricate bias-variance tradeoff.

\paragraph{Comparing HPS to other estimators.}
    \begin{align}
    L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
    &+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2   \lambda^2 \bigtr{\Sigma^{(2)} [(X^{(1)})^\top X^{(1)}]^{-1}  }\nonumber\\
    & + \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  (1-\lambda)^2 \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_AVE_simple}
    \end{align}
    and
    \begin{align}
    L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_STL_simple}
    \end{align}
    %Moreover, there exists a constant $C_0>0$ such that with high probability,
    %\be\label{loss_large}
    %g(a) > g(0) \quad \text{for all $a$ such that }|a|\ge C_0.
    %\ee
    %with high probability for any small constant $\e>0$.

\subsection{Covariate and model shift}
